### 2019-10-16 README.Rmd
added doi and updated title

---
output: github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# Evaluating the Mechanisms for Missing Data in the Enhanced Tuberculosis Surveillance System

[![badge](https://img.shields.io/badge/Launch-Analysis-lightblue.svg)](https://mybinder.org/v2/gh/seabbs/ETSMissing/master?urlpath=rstudio)
[![Documentation](https://img.shields.io/badge/Documentation-click here!-lightgrey.svg?style=flat)](https://www.samabbott.co.uk/ETSMissing)
[![DOI](https://zenodo.org/badge/214470301.svg)](https://zenodo.org/badge/latestdoi/214470301)

[Sam Abbott](https://www.samabbott.co.uk), Hannah Christensen, Ellen Brooks-Pollock

## Background

The Enhanced Tuberculosis Surveillance (ETS) system is a routine surveillance system - with a similar structure to other such systems - that collects data on all notified tuberculosis (TB) cases in England. It is routinely used to study the epidemiology of TB. Routine data often has a large amount of missing data which may not be fully accounted for when used in analyses. This study explores the evidence for associations between missingness in several key outcomes and demographic variables. Any such associations may introduce bias if not accounted for.

## Methods

* Introduce ETS
* Data extraction and management
* Structure of the ETS
* Data completeness
* Drivers of variable completeness (regression)

## Results *Copy from bottom*

* Missing structure
* Drivers of variable completeness

## Conclusions

* Surveillance data is likely to have a high degree of misising data. In the ETS missing for key outcomes is associated with demographic factors such as....
* To avoid biasing analysis studies should make use of imputed data - rather than complete case analysis - and extend their imputation models to other demographic variables that may not be included in the analysis model. 
* This analysis should be repeated in other datasets - for this reason the code is available as an R package.

## Reproducibility

### Repository structure

The repository is structured as an R package. It has the following structure:

- `data-raw`: Raw data processing.
- `data`: Processed data.
- `R`: Supporting R functions.
- `docs:` Documentation for R code.
- `vignettes`: Analysis paper, results, and analysis plan.
- `peer-review`: Documentation required for peer review.

### Manual install

- Install R (analysis run with `3.6.1`) and Rstudio (alternatively use Docker as outlined below).

- Download the analysis folder from [https://github.com/seabbs/ETSMissing/archive/master.zip](https://github.com/seabbs/ETSMissing/archive/master.zip) or use `git clone`, as follows, in the command line (not the R terminal).

```{bash clone-package, eval = FALSE}
git clone https://github.com/seabbs/ETSMissing.git
```

- Once this has been downloaded click on the project file (`ETSMissing.Rproj`).

- Install the analysis dependencies and build the package using the following. To enable more robust reproducibility consider using the [`checkpoint`](https://cran.r-project.org/web/packages/checkpoint/index.html) package versioned locked to R `3.6.1`.

```{r get-packages, eval = FALSE}
#install.packages("devtools")
# To build locally
devtools::install_dev_deps(dependencies = TRUE)
devtools::install()
# Alternatively to remote install
devtools::install_github("seabbs/ETSMissing", dependencies = TRUE)
```

- Load the analysis results by running `vignettes/paper.Rmd`. Alternatively the complete analysis (along with documentation) can be reconstructed using `make` in the project root directory.

- See `data-raw` for data processing and the documentation for implementation details.

### Docker

This analysis was developed in a docker container based on the tidyverse docker image. To run the docker image run:

```{bash, eval = FALSE}
docker run -d -p 8787:8787 --name etsmissing -e USER=etsmissing -e PASSWORD=etsmissing seabbs/etsmissing
```

The rstudio client can be found on port :8787 at your local machines ip. The default username:password is assessbcgpolicychange:assessbcgpolicychange, set the user with -e USER=username, and the password with - e PASSWORD=newpasswordhere. The default is to save the analysis files into the user directory.

If you have access to the required underlying raw data (see [`tbinenglanddataclean`](https://www.samabbott.co.uk/tbinenglanddataclean/)) then the entire analysis can be reproduced from scratch by adding the following to the `docker run` command, with the data saved into `data/tb_data`. The data requirements, and structure, can be found [here](https://www.samabbott.co.uk/tbinenglanddataclean/).

```{bash, eval = FALSE}
--mount type=bind,source=$(pwd)/data/tb_data,target=/home/ETSMissing/data/tb_data
```

Alternatively the analysis environment can be accessed via [binder](https://mybinder.org/v2/gh/seabbs/AssessBCGPolicyChange/master?urlpath=rstudio).